---
title: "Simulation data visulitaion"
output: html_notebook
params: args:myrg
---

1. Install packages 
```{r include=FALSE}
library("SpatialExperiment")
library(ggplot2)
library(gridExtra)
library(dbplyr)
```
### Set 1
## Toy Data seed 59
2. Load in data 
```{r label = "loading set 1 seed 59"}
count = read.csv("set1/toy59/Counts_reformated.csv", sep=",")
comp = read.csv("set1/toy59/synthetic_ST_seed59_1_composition.csv", sep = ",")
geneNamesKey = data.frame("KEY" = 1:dim(count)[1], "Value"=  count[,1])
 #spe = SpatialExperiment(assays=count[,2:101], rowData = geneNames)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

spe - class spatial experiment 

3. Data exploration
```{r }
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```

```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```

```{r include=FALSE}
df1 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df1[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df1[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df1[,i] = 0 }
} 
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 507
2. Load in data 
```{r lable = "Loading set 1 seed 507"}
count = read.csv("set1/toy507/Counts_reformated.csv", sep=",")
comp = read.csv("set1/toy507/synthetic_ST_seed507_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```

```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```

```{r include=FALSE}
df2 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df2[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df2[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df2[,i] = 0 }
} 
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 749

2. Load in data 
```{r lable="Loading set 1 seed 749"}
count = read.csv("set1/toy749/Counts_reformated.csv", sep=",")
comp = read.csv("set1/toy749/synthetic_ST_seed749_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```

```{r include=FALSE}
df3 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df3[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df3[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df3[,i] = 0 }
} 
```

```{r eval=FALSE, include=FALSE, label= "Heatmaps set 1"}
b=seq(0,25875,1725)
for(i in 1:dim(comp)[1]){
x = data.frame("X" =c(rep("s1",dim(df1)[1]),
                            rep("s2",dim(df2)[1]),
                            rep("s3",dim(df3)[1])),
                     "Y"= c(geneNamesKey$KEY,geneNamesKey$KEY, geneNamesKey$KEY),
                      "Fill" = c(df1[,i],
                                 df2[,i],
                                 df3[,i]))
print(paste("Cell type: ", comp$X[i]))
totals = c(as.character(sum(df1[,i])),as.character(sum(df2[,i])),
           as.character(sum(df3[,i])))
ggplot(x, aes(x=X, y=Y))+
  geom_tile(aes(fill=Fill))+
  scale_fill_gradient(low="white",high="Blue")+
  coord_cartesian(clip = "off")+
  #remove x and y axis labels
  labs(x="", y="")+
  #remove extra space
  scale_y_continuous(breaks=b,expand=c(0, 0))+
  #define new breaks on x-axis
  scale_x_discrete(expand=c(0, 0))+
  #set a base size for all fonts
  theme_grey(base_size=8)+
  #theme options
  theme(
    #bold font for legend text
    legend.text=element_text(face="bold"),
    #set thickness of axis ticks
    axis.ticks=element_line(size=0.4),
    #remove plot background
    plot.background=element_blank(),
    #remove plot border
    panel.border=element_blank()
  ) + coord_cartesian(clip = "off")+
  annotate("text",x=1,y=-1,label=totals[1])+
  annotate("text",x=2,y=-1,label=totals[2])+
  annotate("text",x=3,y=-1,label=totals[3])

}

```


Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```


### Set 2
## Toy Data seed 416
2. Load in data 
```{r lable = "loading set 2 seed 416"}
count = read.csv("set2/toy416/Counts_reformated.csv", sep=",")
comp = read.csv("set2/toy416/synthetic_ST_seed416_1_composition.csv", sep = ",")
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r include=FALSE}
df4 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df4[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df4[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df4[,i] = 0 }
} 
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 459
2. Load in data 
```{r lable = "loading set 2 seed 459"}
count = read.csv("set2/toy459/Counts_reformated.csv", sep=",")
comp = read.csv("set2/toy459/synthetic_ST_seed459_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
df5 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df5[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df5[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df5[,i] = 0 }
} 
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 879

2. Load in data 
```{r lable = "Loading set 2 seed 879"}
count = read.csv("set2/toy879/Counts_reformated.csv", sep=",")
comp = read.csv("set2/toy879/synthetic_ST_seed879_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r include=FALSE}
df6 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df6[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df6[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df6[,i] = 0 }
} 
```

```{r eval=FALSE, include=FALSE, label= "Heatmaps set 2"}
plotList = vector("list",dim(comp)[1])
for(i in 1:dim(comp)[1]){
x = data.frame("X" =c(rep("s1",dim(df1)[1]),
                            rep("s2",dim(df2)[1]),
                            rep("s3",dim(df3)[1])),
                     "Y"= c(geneNamesKey$KEY,geneNamesKey$KEY, geneNamesKey$KEY),
                      "Fill" = c(df1[,i],df2[,i],df3[,i]))
print(paste("Cell type: ", comp$X[i]))
plotList[[i]] = ggplot(x, aes(X, Y, fill= Fill)) + 
  geom_tile()+
  scale_fill_gradient(low="white", high="blue")
}
glist = lapply(plotList, ggplotGrob)
print(marrangeGrob(glist, nrow=4, ncol = 3))
```

Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

### Set 3  
## Toy Data seed 108
2. Load in data 
```{r lable = "Loading set 3 seed 108"}
count = read.csv("set3/toy108/Counts_reformated.csv", sep=",")
comp = read.csv("set3/toy108/synthetic_ST_seed108_1_composition.csv", sep = ",")
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```

```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r include=FALSE}
df7 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df7[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df7[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df7[,i] = 0 }
} 
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 411
2. Load in data 
```{r lable = "Loading set 3 seed 411"}
count = read.csv("set3/toy411/Counts_reformated.csv", sep=",")
comp = read.csv("set3/toy411/synthetic_ST_seed411_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
df8 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df8[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df8[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df8[,i] = 0 }
} 
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 598

2. Load in data 
```{r lable = "Loading set 3 seed 598"}
count = read.csv("set3/toy598/Counts_reformated.csv", sep=",")
comp = read.csv("set3/toy598/synthetic_ST_seed598_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```

```{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type", xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```

```{r include=FALSE}
df9 = data.frame(matrix(0,nrow=length(count[,1]),ncol = length(comp[,1])))
for (i in 1:length(comp[,1])) {
  t = comp[i,2:101]
  x = which(t!=0)
  print(comp[i,1]) 
  if(length(x) > 1){
   genes = count[,x+1]
   df9[,i]= rowSums(genes)
   } else if(length(x) ==1) {
   df9[,i]= count[,x+1]
   } else {
   print("Counts zero")
   df9[,i] = 0 }
} 
```

```{r eval=FALSE, include=FALSE, label="Heatmaps set 3"}
plotList = vector("list",dim(comp)[1])
for(i in 1:dim(comp)[1]){
x = data.frame("X" =c(rep("s1",dim(df1)[1]),
                            rep("s2",dim(df2)[1]),
                            rep("s3",dim(df3)[1])),
                     "Y"= c(geneNamesKey$KEY,geneNamesKey$KEY, geneNamesKey$KEY),
                      "Fill" = c(df1[,i],df2[,i],df3[,i]))
print(paste("Cell type: ", comp[i,1]))
plotList[[i]] = ggplot(x, aes(X, Y, fill= Fill)) + 
  geom_tile()+
  scale_fill_gradient(low="white", high="blue")
}
glist = lapply(plotList, ggplotGrob)
print(marrangeGrob(glist, nrow=4, ncol = 3))
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

Create 12 heat maps- on per cell type with the different seeds on the X axis 
```{r}
for(i in 1:dim(comp)[1]){
  x = as.data.frame(cbind(df1[,i],df2[,i],df3[,i],
                          df4[,i],df5[,i],df6[,i],
                          df7[,i],df8[,i],df9[,i]))
  x$sums = rowSums(x)
  table(x$sums > 0)
  x =x[x$sums > 0,]
  gKeys = geneNamesKey[geneNamesKey$KEY %in% row.names(x),]
  y = data.frame("X" =c(rep("s1s1",dim(x)[1]),
                            rep("s1s2",dim(x)[1]),
                            rep("s1s3",dim(x)[1]),
                      rep("s2s1",dim(x)[1]),
                            rep("s2s2",dim(x)[1]),
                            rep("s2s3",dim(x)[1]),
                      rep("s3s1",dim(x)[1]),
                            rep("s3s2",dim(x)[1]),
                            rep("s3s3",dim(x)[1])),
                     "Y"= c(rep(gKeys$KEY,9)),
                      "Fill" = c(x[,1],x[,2],x[,i],
                                 x[,4],x[,5],x[,i],
                                 x[,7],x[,8],x[,i]))
totals=colSums(x[,1:9])
print(paste("Cell type: ", comp[i,1]))
print(ggplot(y, aes(x=X, y=Y))+
  geom_tile(aes(fill=Fill))+
  scale_fill_gradient(low="white",high="Blue")+
  #remove x and y axis labels
  labs(x="", y="")+
  #remove extra space
  scale_y_continuous(expand=c(0, 0))+
  #define new breaks on x-axis
  scale_x_discrete(expand=c(0, 0))+
  #set a base size for all fonts
  theme_grey(base_size=8)+
  #theme options
  theme(
    #bold font for legend text
    legend.text=element_text(face="bold"),
    #set thickness of axis ticks
    axis.ticks=element_line(size=0.4),
    #remove plot background
    plot.background=element_blank(),
    #remove plot border
    panel.border=element_blank()
  ))
  print(totals)
}
```