---
title: "Simulation data visulitaion"
output: html_notebook
params: args:myrg
---

1. Install packages 
```{r,echo=FALSE}
library("SpatialExperiment")
```
### Set 1
## Toy Data seed 59
2. Load in data 
```{r label = "loading set 1 seed 59"}
count = read.csv("set1/toy59/Counts_reformated.csv", sep=",")
comp = read.csv("set1/toy59/synthetic_ST_seed59_1_composition.csv", sep = ",")
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 507
2. Load in data 
```{r lable = "Loading set 1 seed 507"}
count = read.csv("set1/toy507/Counts_reformated.csv", sep=",")
comp = read.csv("set1/toy507/synthetic_ST_seed507_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 749

2. Load in data 
```{r lable="Loading set 1 seed 749}
count = read.csv("set1/toy749/Counts_reformated.csv", sep=",")
comp = read.csv("set1/toy749/synthetic_ST_seed749_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```
### Set 2
## Toy Data seed 416
2. Load in data 
```{r lable = "loading set 2 seed 416}
count = read.csv("set2/toy416/Counts_reformated.csv", sep=",")
comp = read.csv("set2/toy416/synthetic_ST_seed416_1_composition.csv", sep = ",")
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 459
2. Load in data 
```{r lable = "loading set 2 seed 459"}
count = read.csv("set2/toy459/Counts_reformated.csv", sep=",")
comp = read.csv("set2/toy459/synthetic_ST_seed459_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 879

2. Load in data 
```{r lable = "Loading set 2 seed 879"}
count = read.csv("set2/toy879/Counts_reformated.csv", sep=",")
comp = read.csv("set2/toy879/synthetic_ST_seed879_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

### Set 3  
## Toy Data seed 108
2. Load in data 
```{r lable = "Loading set 3 seed 108"}
count = read.csv("set3/toy108/Counts_reformated.csv", sep=",")
comp = read.csv("set3/toy108/synthetic_ST_seed108_1_composition.csv", sep = ",")
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 411
2. Load in data 
```{r lable = "Loading set 3 seed 411"}
count = read.csv("set3/toy411/Counts_reformated.csv", sep=",")
comp = read.csv("set3/toy411/synthetic_ST_seed411_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 598

2. Load in data 
```{r lable = "Loading set 3 seed 598"}
count = read.csv("set3/toy598/Counts_reformated.csv", sep=",")
comp = read.csv("set3/toy598/synthetic_ST_seed598_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
```
``{r}
plot(count_sums,main="Scatter of summed expression per spot", xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,main="Scatter of summed cell types per spot", xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, main="Scatter of counts per cell type" xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,main="Scatter of counts per gene", xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums, main="Distribution of summed counter per spot")
hist(spot_sums, main="Distribution of summed cell types per spot")
hist(cT_sums,main="Distribution of counter per cell type")
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```
