---
title: "Simulation data visulitaion"
output: html_notebook
---

1. Install packages 
```{r}
library("SpatialExperiment")
```

## Toy Data seed 59
2. Load in data 
```{r}
count = read.csv2("toy59/Counts_reformated.csv", sep=",")
comp = read.csv("toy59/s.3/synthetic_ST_seed59_1_composition.csv", sep = ",")
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
plot(count_sums,xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```
```{r}
hist(count_sums)
hist(spot_sums)
hist(cT_sums)
```
Explore cell type more
```{r}
summary(cT_sums)
comp$X[which(cT_sums==0)+1]
comp$X[which(cT_sums >= 200)+1]
```

## Toy Data seed 507
2. Load in data 
```{r}
count = read.csv2("toy507/Counts_reformated.csv", sep=",")
comp = read.csv("toy507/s.3/synthetic_ST_seed507_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
plot(count_sums,xlab="Spots", ylab="Gene counts per spot (sum)")
```

```{r}
plot(spot_sums,xlab="Spots", ylab="Cell Type counts per spot (sum)")
```

```{r}
plot(cT_sums, xlab = "Cell type", ylab="Total counts per cell Type")
```

```{r}
plot(gene_sums,xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```


```{r}
hist(count_sums)
hist(spot_sums)
hist(cT_sums)
```

## Toy Data seed 749

2. Load in data 
```{r}
count = read.csv2("toy749/Counts_reformated.csv", sep=",")
comp = read.csv("toy749/s.3/synthetic_ST_seed749_1_composition.csv", sep = ",")
dim(count)
dim(comp)
```
About the data... 
c is the file containing count information with each row being genes (25,874) and each column spots (100)

comp is the file containing cell type information with each row being a cell type (11) and each column the number per spot (100)

3. Data exploration
```{r}
count_sums = colSums(count[2:25875,2:101])
gene_sums = rowSums(count[2:25875,2:101])
spot_sums = colSums(comp[2:12,2:101])
cT_sums = rowSums(comp[2:12,2:101])
plot(count_sums,xlab="Spots", ylab="Gene counts per spot (sum)")
plot(spot_sums,xlab="Spots", ylab="Cell Type counts per spot (sum)")
plot(cT_sums, xlab = "Cell type", ylab="Total counts per cell Type")
plot(gene_sums,xlab="Genes", ylab="Total counts per gene")
abline(h=mean(gene_sums),col="red")
```


```{r}
hist(count_sums)
hist(spot_sums)
hist(cT_sums)
```